Title

### 2014-11-14 manuscript.tex
Moved to Geophysics again. Will focus on code
Moving the focus of the paper to the algorithm and implementation. This
will fit nicely on the Software section of Geophysics.

\documentclass[extra,mreferee]{gji}
%\documentclass[extra,paper]{gji}
\usepackage{timet}
\usepackage{graphicx}

\title[Quantifying the error in tesseroid forward modeling]{
Quantifying the error in forward modeling gravitational fields using tesseroids
with adaptive discretization
}

\author[Uieda et al.]{
Leonardo Uieda$^{1,2}$,
Val\'eria C. F. Barbosa$^{2}$,
Vanderlei C. Oliveira Jr$^{2}$,
Carla Braitenberg$^{3}$
\\
$^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil. \\
$^2$Observat\'orio Nacional, Rio de Janeiro, Brazil. \\
$^3$Department of Mathematics and Geosciences, University of Trieste, Trieste, Italy.
}

@@

\label{firstpage}
\maketitle

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

@@

\citet{Grombein2013} follows \citet{Heck2007} and solves the integrals by a
Taylor expansion around the geometric center of the tesseroid.

We will follow \citet{Asgharzadeh2007} and perform the numerical integration
using the Gauss-Legendre Quadrature (GLQ) instead.

@@

$N$ is the number of points used in the GLQ, i.e. the order of the quadrature.

@@

\section{Adaptive discretization algorithm}

The quantity of, and therefore distance between, point masses regulates the
accuracy of the numerical integration \citep{Ku1977}.

@@

One way to do this would be increase the order of the quadrature $N$ in all
three dimensions.

An alternative way would be to fix the quadrature order (e.g., $N=2$) and
instead divide the tesseroid in half along each dimension.

@@

\citet{Li2011} developed an algorithm based on the tesseroid division approach.

In their algorithm, a tesseroid is divided in half along one of its dimensions
if the minimum distance to the computation point is smaller than a multiple of
the size of the tesseroid along that dimension.

The criterion check and (possible) subsequent division is repeated recursively
(i.e., to each of the smaller tesseroids and so on).

The recursion stops when none of the smaller tesseroids fail the criterion.

@@

@@

\bibliographystyle{gji}
\bibliography{references.bib}
\end{document}

>>>>

\documentclass[paper,twocolumn]{geophysics}
%\documentclass[manuscript]{geophysics}
%\documentclass[manuscript,endfloat]{geophysics}
\usepackage{graphicx}
\newcommand{\mitbf}[1]{\mathbf{{#1}}}

@@

\title{Title}
\lefthead{Meh}
\righthead{Meh}
% manuscript number
\ms{Submission}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\address{
\footnotemark[1] Universidade do Estado do Rio de Janeiro,
Rio de Janeiro, Brazil.
\\
\footnotemark[2] Observat\'orio Nacional,
Rio de Janeiro, Brazil.
\\
\footnotemark[3] Department of Mathematics and Geosciences,
University of Trieste, Trieste, Italy.
}

\author{
    Leonardo Uieda\footnotemark[1]\footnotemark[2]
    Val\'eria C. F. Barbosa\footnotemark[2],
    Vanderlei C. Oliveira Jr\footnotemark[2],
    and
    Carla Braitenberg\footnotemark[3]
}

\maketitle

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}

@@

We will follow \citet{Asgharzadeh2007} and perform the numerical integration
using the Gauss-Legendre Quadrature (GLQ).

@@

$N$ is the order of the quadrature, i.e. the number of points used in the GLQ.

@@

The accuracy of the integration
depends on the number of point masses used in the summation.

\citet{Ku1977} showed that it also depends on the ratio between
the distance to the computation point and the distance between adjacent nodes.

Figure~\ref{fig:glqerrorsample}
illustrates this effect on the $g_{xy}$ gravity gradient component.

The $g_{xy}$ component was produced by a
$7^\circ \times 7^\circ \times 20\ km$ tesseroid
with $2.67\ g.cm^{-3}$ density
and top at $z=0\ km$.

The maps were calculated on a regular grid
with $100\times100$ points.

The GLQ order used was the same
in the radial, longitudinal, and latitudinal dimensions.

Figure~\ref{fig:glqerrorsample}a shows the $g_{xy}$ component
calculated at 400 km height using
GLQ with order two (two point masses in the longitude, latitude, and vertical
directions, respectively).

Figure~\ref{fig:glqerrorsample}b shows $g_{xy}$ computed with order two
GLQ as well but at 150 km height.

Notice that the computed effect is concentrated around each point mass (black
dots) and does not resemble the effect of a tesseroid.

@@

One way to accomplish this would be increase the order of the quadrature
$N$ in all three directions.

Figure~\ref{fig:glqerrorsample}c shows the $g_{xy}$ component calculated at
150km height but with a GLQ order of 30.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
    \caption{
        Example of the effect of varying
        the computation height
        and the number of point masses in the Gauss-Legendre Quadrature
        (i.e., the GLQ order).
        Black circles represent the horizontal location of the GLQ nodes
        (not shown for GLQ order 30).
        a) Field calculated at $400\ km$ height using GLQ order 2.
        The effect is similar to that presented by \citet{Asgharzadeh2007}.
        b) At $150\ km$ height and order 2,
        the field resembles that of
        four point sources located at the GLQ nodes.
        This effect was shown by \citet{Ku1977}
        for the case of the $g_z$ component of a right rectangular prism.
        c) At $150\ km$ but with a higher GLQ order of 30,
        the field is similar to that expected for a single mass source.
    }
    \label{fig:glqerrorsample}
\end{figure}

\subsection{Adaptive discretization}

\citet{Li2011} proposed an alternative way to decrease the distance between
point masses and achieve an accurate integration.

Instead of increasing the GLQ order,
they divide the tesseroid into smaller tesseroids while keeping the number of
point masses per tesseroid fixed.

@@

The criterion for dividing a tesseroid is that
the distance to the computation point must be smaller than
a constant times the size of the tesseroid.

The size of the tesseroid is used as proxy for the distance between point
masses.

This procedure is repeated recursively
(i.e., to each of the smaller tesseroids)
until all tesseroids are within the acceptable distance-size ratio
or a minimum size is achieved.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical evaluation of the accuracy}

@@

\bibliographystyle{seg}
\bibliography{references.bib}
\end{document}
